---
title: "School Performance"
author: "Advait"
date: "4/5/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imported the dataset, and viewed the summary and structure

```{r school}
file <- file.choose()
school <- read.csv(file, sep = ",", header = TRUE, stringsAsFactors = FALSE)
View(school)

colnames(school)
colnames(school) <- c("School", "Section", "Very Ahead", "Middling", "Behind", "More Behind", "Very Behind", "Completed")

summary(school) # Provides information about the spread of the data within the columns
str(school) # school is "character" and others are integer type
table(school$School) # There are 5 schools. 
table(school$Section) # There are 13 sections. 

```

## Visualization 1:

```{r visualization 1}
count3 <- table(school$School, school$Section)
plot3 <- barplot(count3, col=c("lightblue4","lightblue3","lightblue2","lightblue1","lightcyan")
                 , ylab = "Schools", xlab = "Sections"
                 , main = "Number of Schools across Sections"
                 , legend=rownames(count3))
```
# There are 13 sections in A, 12 sections in B, 3 sections in C, and 1 section in D and E respectively

# Student-level wise distributions:

```{r visualization 2}
# Very Ahead: Histogram with mean line

hist1 <- hist(school$`Very Ahead`
              , col = "salmon", xlab = "Very Ahead", main = "Very Ahead: All Schools"
              , xlim = c(0,60))
abline(v=mean(school$`Very Ahead`),col="blue",lwd=4)
```

```{r visualization 3}
# Middling: Histogram with mean line

hist2 <- hist(school$Middling, col="indianred1", xlab = "Middling", main = "Middling: All Schools"
              , xlim = c(0,60))
abline(v=mean(school$Middling),col="blue",lwd=4)
```

```{r visualization 4}
# Behind: Histogram with mean line

hist3 <- hist(school$Behind, col="indianred2", xlab = "Behind", main = "Behind: All Schools", xlim = c(0,60))
abline(v=mean(school$Behind),col="blue",lwd=4)
```

```{r visualization 5}
# More Behind: Histogram with mean line

hist4 <- hist(school$`More Behind`, col="indianred3", xlab = "More Behind", main = "More Behind: All Schools"
              , xlim = c(0,60))
abline(v=mean(school$`More Behind`),col="blue",lwd=4)
```

```{r visualization 6}
# Very Behind: Histogram with mean line

hist5 <- hist(school$`Very Behind`, col="indianred4", xlab = "Very Behind", main = "Very Behind: All Schools"
              , xlim = c(0,60))
abline(v=mean(school$`Very Behind`),col="blue",lwd=4)
```

```{r visualization 7}
# Completed: Histogram with mean line

hist6 <- hist(school$Completed, col="red4", xlab = "Completed", main = "Completed: All Schools"
              , xlim = c(0,60))
abline(v=mean(school$Completed),col="blue",lwd=4)
```

# Conclusion through visualizations:

# There are no students who are very ahead in the course, in any school.
# Most of the students are 1-5 lessons behind, across all the schools.
# Following is the order of sizes of levels of completion:
#             Behind > Completed > Middling > Very Behind > More Behind > Very Ahead 

# Schoolwise distributions:

```{r visualization 8}
# Very Ahead: Schoolwise bar-chart

Very_ahead_schoolwise <- tapply(school$`Very Ahead`, list(school$School), FUN = sum)
bar1 <- barplot(Very_ahead_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "Very Ahead: Schoolwise", col = "green3")
```

```{r visualization 9}
# Middling: Schoolwise bar-chart

Middling_schoolwise <- tapply(school$Middling, list(school$School), FUN = sum)
bar2 <- barplot(Middling_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "Middling: Schoolwise", col = "olivedrab3")
```

```{r visualization 10}
# Behind: Schoolwise bar-chart

Behind_schoolwise <- tapply(school$Behind, list(school$School), FUN = sum)
bar3 <- barplot(Behind_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "Behind: Schoolwise", col = "darkolivegreen1")
```

```{r visualization 11}
# More Behind: Schoolwise bar-chart

More_behind_schoolwise <- tapply(school$`More Behind`, list(school$School), FUN = sum)
bar4 <- barplot(More_behind_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "More Behind: Schoolwise", col = "darkolivegreen2")

```

```{r visualization 12}
# Very Behind: Schoolwise bar-chart

Very_behind_schoolwise <- tapply(school$`Very Behind`, list(school$School), FUN = sum)
bar5 <- barplot(Very_behind_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "Very Behind: Schoolwise", col = "darkolivegreen3")

```

```{r visualization 13}
# Completed: Schoolwise bar-chart

Completed_schoolwise <- tapply(school$Completed, list(school$School), FUN = sum)
bar6 <- barplot(Completed_schoolwise, xlab="Schools", ylab="Number of Students", ylim=c(0,400)
                , main = "Completed: Schoolwise", col = "darkolivegreen4")
```

# Conclusions:
# The number of students across schools are: 
#        A > B > C > E > D
# The number of students who are behind (1-5 lessons behind) are strickingly higher than the other levels of completion.

```{r Counts}
# Adding count of total students:
school$`Very Ahead` <- as.numeric(school$`Very Ahead`)
school$Middling <- as.numeric(school$Middling)
school$Behind <- as.numeric(school$Behind)
school$`More Behind` <- as.numeric(school$`More Behind`)
school$`Very Behind` <- as.numeric(school$`Very Behind`)
school$Completed <- as.numeric(school$Completed)

school$Total_students <- school$`Very Ahead` + school$Middling + school$Behind + school$`More Behind` +school$`Very Behind` + school$Completed

```

```{r Subsetting}
SchoolA <- subset(school,school$School == "A") 
SchoolB <- subset(school,school$School == "B") 
SchoolC <- subset(school,school$School == "C") 
SchoolD <- subset(school,school$School == "D") 
SchoolE <- subset(school,school$School == "E")
```

```{r Distribution A}
Distribution_A <- c(sum(SchoolA$`Very Ahead`)
                    ,sum(SchoolA$Middling)
                    ,sum(SchoolA$Behind)
                    ,sum(SchoolA$`More Behind`)
                    ,sum(SchoolA$`Very Behind`)
                    ,sum(SchoolA$Completed))

Distribution_A_percent <- c(sum(SchoolA$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolA$Middling)/sum(school$Total_students)
                    ,sum(SchoolA$Behind)/sum(school$Total_students)
                    ,sum(SchoolA$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolA$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolA$Completed)/sum(school$Total_students))
```

```{r Distribution B}
Distribution_B <- c(sum(SchoolB$`Very Ahead`)
                    ,sum(SchoolB$Middling)
                    ,sum(SchoolB$Behind)
                    ,sum(SchoolB$`More Behind`)
                    ,sum(SchoolB$`Very Behind`)
                    ,sum(SchoolB$Completed))

Distribution_B_percent <- c(sum(SchoolB$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolB$Middling)/sum(school$Total_students)
                    ,sum(SchoolB$Behind)/sum(school$Total_students)
                    ,sum(SchoolB$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolB$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolB$Completed)/sum(school$Total_students))
```

```{r Distribution C}
Distribution_C <- c(sum(SchoolC$`Very Ahead`)
                    ,sum(SchoolC$Middling)
                    ,sum(SchoolC$Behind)
                    ,sum(SchoolC$`More Behind`)
                    ,sum(SchoolC$`Very Behind`)
                    ,sum(SchoolC$Completed))

Distribution_C_percent <- c(sum(SchoolC$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolC$Middling)/sum(school$Total_students)
                    ,sum(SchoolC$Behind)/sum(school$Total_students)
                    ,sum(SchoolC$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolC$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolC$Completed)/sum(school$Total_students))
```

```{r Distribution D}
Distribution_D <- c(sum(SchoolD$`Very Ahead`)
                    ,sum(SchoolD$Middling)
                    ,sum(SchoolD$Behind)
                    ,sum(SchoolD$`More Behind`)
                    ,sum(SchoolD$`Very Behind`)
                    ,sum(SchoolD$Completed))

Distribution_D_percent <- c(sum(SchoolD$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolD$Middling)/sum(school$Total_students)
                    ,sum(SchoolD$Behind)/sum(school$Total_students)
                    ,sum(SchoolD$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolD$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolD$Completed)/sum(school$Total_students))
```

```{r Distribution E}
Distribution_E <- c(sum(SchoolE$`Very Ahead`)
                    ,sum(SchoolE$Middling)
                    ,sum(SchoolE$Behind)
                    ,sum(SchoolE$`More Behind`)
                    ,sum(SchoolE$`Very Behind`)
                    ,sum(SchoolE$Completed))

Distribution_E_percent <- c(sum(SchoolE$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolE$Middling)/sum(school$Total_students)
                    ,sum(SchoolE$Behind)/sum(school$Total_students)
                    ,sum(SchoolE$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolE$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolE$Completed)/sum(school$Total_students))
```

```{r Distribution E}
Distribution_E <- c(sum(SchoolE$`Very Ahead`)
                    ,sum(SchoolE$Middling)
                    ,sum(SchoolE$Behind)
                    ,sum(SchoolE$`More Behind`)
                    ,sum(SchoolE$`Very Behind`)
                    ,sum(SchoolE$Completed))

Distribution_E_percent <- c(sum(SchoolE$`Very Ahead`)/sum(school$Total_students)
                    ,sum(SchoolE$Middling)/sum(school$Total_students)
                    ,sum(SchoolE$Behind)/sum(school$Total_students)
                    ,sum(SchoolE$`More Behind`)/sum(school$Total_students)
                    ,sum(SchoolE$`Very Behind`)/sum(school$Total_students)
                    ,sum(SchoolE$Completed)/sum(school$Total_students))
```

```{r Binding Distributions}
Distribution_Schoolwise <- rbind(Distribution_A,Distribution_B,Distribution_C,Distribution_D,Distribution_E)
Distribution_Schoolwise_percent <- rbind(Distribution_A_percent,Distribution_B_percent,Distribution_C_percent
                                        ,Distribution_D_percent,Distribution_E_percent)
```

```{r Pie charts}
par(mfrow=c(3,2))
pieA <- pie(Distribution_A, labels = c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                                , "Completed"), main = "School A: Course Status")

pieB <- pie(Distribution_B, labels = c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed"), main = "School B: Course Status")

pieC <- pie(Distribution_C, labels = c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed"), main = "School C: Course Status")

pieD <- pie(Distribution_D, labels = c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed"), main = "School D: Course Status")

pieE <- pie(Distribution_E, labels = c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed"), main = "School E: Course Status")
```

```{r Distribution Schoolwise}
Distribution_Schoolwise <- as.data.frame(Distribution_Schoolwise)
colnames(Distribution_Schoolwise) <- c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed")
Distribution_Schoolwise_percent <- as.data.frame(Distribution_Schoolwise_percent)
colnames(Distribution_Schoolwise_percent) <- c("Very Ahead", "Middling","Behind","More Behind", "Very Behind"
                                       , "Completed")
Distribution_Schoolwise_percent <- 100*Distribution_Schoolwise_percent

Distribution_Schoolwise
Distribution_Schoolwise_percent
```

# Percentage split within each school for course completion levels:

# Very Ahead: 0 students
# Middling: School B > School D > School C > School A > School E
# Behind: School A > School E > School C > School B > School D
# More Behind: School D > School A > School E > School C > School B
# Very Behind: School D > School A > School C > School E > School B
# Completed: School B > School E > School C > School A > School D

# Number of students across course completion levels within each school:
# Very Ahead: 0 students
# Middling: School B > School A > School C = School E > School D
# Behind: School A > School B > School E > School C > School D
# More Behind: School A > School B = School C = School D = School E
# Very Behind: School A > School C = School D = School E > School B
# Completed: School B > School A > School E > School C > School D

# S
# School A: Poor
# School B: Very Good
# School C: Average
# School D: Very Poor
# School E: Good

# School A needs to encourage students to improve the pace and bring more students to middling level.

# School B is performing really well. Their focus area should be uplifting students to complete the course. 
# Having many students complete the course, it can act as a positive pressure for the other students.

# School C needs to focus on all the areas equally, since it is in the average level in comparison to other 
# schools.

# Counseling might be needed for the students in School D, as there are not many students in this school, 
# which might have a negative impact, which is visible in the performance. 
# It is possible that the students might need academic or other help.

# School E is performing good. It needs to uplift the students in the behind level, 
# and bring them to atleast middling level.
